Light versus Dark in Strong Lensing Galaxies: Dark matter halos are rounder than their stars

### 2014-12-02 shapes/paper.tex
A few changes.

\author[Bruderer]{Claudio Bruderer$^{1}$\thanks{E-mail: claudio.bruderer@phys.ethz.ch}, J. I. Read$^{1,2}$, P. Saha$^{3}$, J. Coles$^{4}$\\
$^{1}$Institute for Astronomy, Department of Physics, ETH Z\"urich, Wolfgang-Pauli-Strasse 27, CH-8093 Z\"urich, Switzerland\\
$^{2}$Department of Physics, University of Surrey, Guildford, GU2 7XH, UK\\
$^{3}$Institute for Theoretical Physics, University of Z\"urich, Winterthurerstrasse 190, 8057 Z\"urich, Switzerland\\
$^{4}$Department of Biology and Health, Versailles Saint-Quentin-en-Yvelines University, France
}

@@

\begin{abstract}
We measure the shape and alignment of stars and dark matter in 11 strong lensing galaxies with good quality data. We find that in all cases the dark matter halos are more elliptical than the light distribution over the range $R_e < R < 5R_e$. As we average over larger radii, the lenses become increasingly elliptical both in their dark matter and stars, but dark matter halos are never more elliptical than $s_{dm} = 1.15$, while their stars can extend to $s_* > 1.4$ (where $s$ is the ratio of the largest to smallest eigenvalue of the 2D moment of inertia tensor). Three systems have very high stellar ellipticity ($s_* > 1.6$) and correspondingly high alignment between light and dark. One of these -- B1608 -- is a known merging pair; we suggest that the other two (B0712 and B2016) may also be recent post-merger systems. Galaxies with high dark matter ellipticity and weak external shear show very strong alignment between light and dark; those with strong shear ($\gamma \simgt 0.1$) can be very highly misaligned. This is reassuring since isolated misaligned galaxies are expected to be unstable.

Our results provide a new constraint on galaxy formation models that must explain the origin of very round dark matter halos (not expected in pure dark matter only simulations), and highly misaligned systems. Such misalignments also present a new challenge for alternative gravity theories in which the light and dark must necessarily be highly correlated.
\end{abstract}

@@

\Glass\ (Gravitational Lensing AnalysiS Software) is a new `non-parametric' lens modelling framework. (Non-parametric here simply means that we deliberately use many more parameters than data constraints such that our system of equations is under-constrained.) \Glass\ shares some aspects
with an earlier code \PixeLens\ \citep{Saha2004,2008ApJ...679...17C}. However,
\Glass\ -- which contains all new code written from the ground up --
significantly improves upon \PixeLens\ in several key ways:

\begin{enumerate}
\item At the heart of \Glass\ is a new uniform sampling algorithm for high
    dimensional spaces \citep{2012MNRAS.425.3077L}. This allows for large
    ensembles of $>10,000$ models to be efficiently generated. 
\item \Glass\ provides a modular framework that allows new priors to be added
    and modified easily.
\item The basis functions approximating a model can be easily changed (in this
    paper, we assume pixels as in \PixeLens). 
\item With so many models in the final ensemble, we can afford to apply
    non-linear constraints (for example stellar kinematic data; or the removal of
    models with spurious extra images) to accept/reject models in a post-processing
    step.
\item The central region of the mass map can have a higher resolution to more
    efficiently capture steep models.
\item Stellar density can be used as an additional constraint on the models. 
\item Point or extended mass objects can be placed in the field.
\end{enumerate}

\Glass\ is described in detail in the code paper: \citet{2014arXiv1401.7990C}. Here, we simple list the default priors we use to model each lens. Where the priors differ from these defaults for any given lens, this is spelled out in data Table \ref{tab:lensmodelling}:

\begin{enumerate} 
\item Observational priors (redshifts, image positions relative to lens position).
\item Image parity is enforced.
\item Reconstructed stellar mass map.
\item Local density gradient is always pointing within 50$^{\circ}$ to the center and the slope is at most 0 everywhere. In other words, the density profile peaks in the center.
\item Fixed cosmology: $\Omega_{m} = 0.28$, $\Omega_{\Lambda} = 0.72$, $\Omega_{k} = 0$, $H_{0}^{-1} = 13.7$ Gyr.
\item A flat prior on the magnitude of the external shear $|\gamma|$ on $[0,1)$.
\end{enumerate}

In case of more information, the following priors can be turned on:

\begin{enumerate}
\item[(vii)] A flat prior on measured time delays within the $\pm 1\sigma$-interval.
\item[(viii)] External point masses to model satellite galaxies outside but close to the Einstein radius.
\item[(ix)] Point symmetry $\kappa(x,y) = \kappa(-x,-y)$ in case there is reason to assume the surface mass profile to be symmetric (3+1 quad configurations and the double \textit{Q0957}).
\end{enumerate}

@@

\begin{table*}
  \begin{center}
    \begin{tabular}{l r r r r l}
      Lens    & \multicolumn{1}{c}{$z_{L}$} & \multicolumn{1}{c}{$z_{S}$} & \multicolumn{1}{c}{$\Delta\theta$ [kpc]} & \multicolumn{1}{c}{$R_{L}$/$R_{e}$} & Environment \\ \hline
      0047 & 0.485 & 3.60 & 12.82 & $1.45\pm0.04$ & G(9) \\
      0414 & 0.960 & 2.64 & 16.01 & $1.85\pm0.05$ & ... \\
      0712 & 0.410 & 1.34 & 6.82  & $1.15\pm0.03$ & ... \\
      0911 & 0.769 & 2.8  & 23.16 & $3.09\pm0.05$ & C \\
      0957 & 0.356 & 1.41 & 29.98 & $3.51\pm0.04$ & C \\
      1115 & 0.310 & 1.72 & 10.76 & $2.94\pm0.06$ & G(13) \\
      1422 & 0.337 & 3.62 & 6.02  & $4.49\pm0.06$ & G(17) \\
      1608 & 0.630 & 1.39 & 13.92 & $1.82\pm0.01$ & !CHECK! \\
      2016 & 1.010 & 3.3  & 26.22 & $6.12\pm0.14$ & C(69) \\
      2045 & 0.870 & 1.28 & 14.46 & $1.48\pm0.03$ & !SEE TEXT! \\
      2237 & 0.039 & 1.7  & 1.40  & $0.89\pm0.01$ & ... \\
    \end{tabular}
    \caption[width=\linewidth]{Table with lens properties \citep[see][]{2011ApJ...740...97L}. References in Section~\ref{sec:data} \texbf{expand}}
    \label{tab:lensproperties}
  \end{center}
\end{table*}

\begin{table*}
  \begin{center}
    \begin{tabular}{l r r r r l l}
      Lens    & \multicolumn{1}{c}{A ['']} & \multicolumn{1}{c}{B ['']} & \multicolumn{1}{c}{C ['']} & \multicolumn{1}{c}{D ['']} & Time delays [days] & Additional priors \\ \hline
      0047 & 1.270 & -0.630 & 0.520 & -0.730 & ... & None \\
           & 0.105 & -0.995 & -1.045 & 0.705 & &  \\
      0414 & -0.472 & -1.061 & -1.1947 & 0.885 & ... & None \\
           & 1.277 & -0.661 & -0.255 & -0.361 & &  \\
      0712 & -0.013 & 0.795 & 0.747 & -0.391 & ... & None \\
           & -0.804 & -0.156 & -0.292 & 0.307 & &  \\
      0712 & -0.013 & 0.795 & 0.747 & -0.391 & ... & None \\
           & -0.804 & -0.156 & -0.292 & 0.307 &  &  \\
      0957$^{a}$ & 1.408 & 0.182 & 2.860 & -1.540 & $\Delta t_{BA}=416.5^{+1}_{-1}$ & Symm \\
           & 5.034 & -1.018 & 3.470 & -0.050 &  &  \\
      1115 & 0.355 & -0.909 & -1.093 & 0.717 & $\Delta t_{BA}=12.0^{+2}_{-2} \ \Delta t_{DC}=4.4^{+3.2}_{-2.4}$ & None \\
           & 1.322 & -0.714 & -0.260 & -0.627 & &  \\
      1422 & 1.079 & 0.357 & 0.742 & -0.205 & ... & Symm \\
           & -0.095 & 0.973 & 0.656 & -0.147 &  &  \\
      1608 & -1.300 & -0.560 & -1.310 & 0.570 & $\Delta t_{BA}=31.5^{+2}_{-1} \ \Delta t_{CA}=36.0^{+1.5}_{-1.5} \ \Delta t_{DA}=77.0^{+2}_{-1}$ & None \\
           & -0.800 & 1.160 & 0.700 & -0.080 &  &  \\
      2016 & -1.735 & 0.335 & 0.437 & 1.268 & ... & None \\
           & 1.778 & -1.450 & -1.435 & 0.276 &  &  \\
      2045 & 1.121 & 1.409 & 1.255 & -0.507 & ... & $\gamma(0.1)$, Symm \\
           & 0.824 & 0.035 & 0.576 & -0.183 &  &  \\
      2237 & 0.598 & -0.075 & 0.791 & -0.710 & ... & None \\
              & 0.758 & -0.939 & -0.411 & 0.271 &  &  \\
    \end{tabular}
    \caption[width=\linewidth]{Table with lens properties relevant for modelling (positions, time delays) \newline $^{a}$ Images A and B are a pair, C and D the other \texbf{expand}}
    \label{tab:lensmodelling}
  \end{center}
\end{table*}

>>>>

\author[Bruderer]{Claudio Bruderer$^{1}$\thanks{E-mail: claudio.bruderer@phys.ethz.ch}, J. I. Read$^{1,2}$, P. Saha$^{3}$, J. Coles$^{4}$\\
$^{1}$Institute for Astronomy, Department of Physics, ETH Z\"urich, Wolfgang-Pauli-Strasse 27, CH-8093 Z\"urich, Switzerland\\
$^{2}$Department of Physics, University of Surrey, Guildford, GU2 7XH, UK\\
$^{3}$Institute for Theoretical Physics, University of Z\"urich, Winterthurerstrasse 190, 8057 Z\"urich, Switzerland\\
$^{4}$Exascale Research Computing Lab, Campus Teratec, 2 Rue de la Piquetterie, 91680 Bruyeres-le-Chatel, France
}

@@

\begin{abstract}
We measure the shape and alignment of stars and dark matter in 11 strong
lensing galaxies and find that in all cases the dark matter halos are rounder
than the light distribution over the range $R_e < R < 5R_e$.  Averaging over
larger radii, the lenses become increasingly elliptical both in their dark
matter and stellar components, as defined using the ratio $s$ of the largest to
smallest eigenvalues of the 2D moment of inertia tensor of the projected
lensing mass.  While dark matter halos are never more elliptical than $s_{dm} =
1.15$, their stars can extend to $s_* > 1.4$ Three systems, in particular, have
very high stellar ellipticity ($s_* > 1.6$) and correspondingly high alignment
between light and dark. One of these -- B1608 -- is a known merging pair; we
suggest that the other two (B0712 and B2016) may also be recent post-merger
systems.  Galaxies with high dark matter ellipticity and weak external shear
show very strong alignment between light and dark; those with strong shear
($\gamma \simgt 0.1$) can be very highly misaligned. This is reassuring since
isolated misaligned galaxies are expected to be unstable.

Our results provide a new constraint on galaxy formation models that must
explain the origin of very round dark matter halos and highly misaligned
systems. Such misalignments also present a new challenge for alternative
gravity theories in which the light and dark must necessarily be highly
correlated.

\end{abstract}

@@

\Glass{} is a new `non-parametric' gravitationl lens modelling framework
\citep{2014arXiv1401.7990C} that is capable of generating a distribution of
mass models given a set of basic assumptions, or priors.  In this context,
non-parametric means that  many more parameters than data constraints are used
such that system of equations describing the lensing mass distribution is
under-constrained.  \Glass{} shares some aspects with an earlier code
\PixeLens{} \citep{Saha2004,2008ApJ...679...17C}, but has a completely
different code base. It key improvements upon \PixeLens{} which are relevant
for this paper include:

\begin{enumerate}
\item A new uniform sampling algorithm for high
    dimensional spaces \citep{2012MNRAS.425.3077L}. This allows for large
    ensembles of $>10,000$ models to be efficiently generated. 
%\item \Glass\ provides a modular framework that allows new priors to be added
%    and modified easily.
%\item The basis functions approximating a model can be easily changed (in this
%    paper, we assume pixels as in \PixeLens). 
\item With so many models in the final ensemble, we can afford to apply
    non-linear constraints (for example stellar kinematic data; or the removal of
    models with spurious extra images) to accept/reject models in a post-processing
    step.
\item The central region of the mass map can have a higher resolution to more
    efficiently capture steep models.
\item Stellar density can be used as an additional lower-bound constraint on the models. 
%\item Point or extended mass objects can be placed in the field.
\end{enumerate}

%\Glass{} is described in detail in the code paper: \citet{2014arXiv1401.7990C}.
In this paper, the default priors used to model each lens are listed below.  Table
\ref{tab:lensmodelling} identifies lenses for which we have modified these defaults.
For each lens we assume:

\begin{enumerate} 
\item Observational priors (redshifts, image positions relative to lens position).
\item Image parity is enforced.
\item A reconstructed stellar mass map.
\item The local density gradient is always pointing within 50$^{\circ}$ to the center and the slope is at most 0 everywhere. In other words, the density profile peaks in the center.
\item The fixed cosmology $\Omega_{m} = 0.28$, $\Omega_{\Lambda} = 0.72$, $\Omega_{k} = 0$, $H_{0}^{-1} = 13.7$ Gyr.
\item A flat prior on the magnitude of the external shear $|\gamma|$ on $[0,1)$.
\end{enumerate}

In cases where more information is available, the following priors are also used:

\begin{enumerate}
\item[(vii)] A flat prior on measured time delays within the $\pm 1\sigma$-interval.

\item[(viii)] External point masses to model satellite galaxies outside but
close to the Einstein radius.

\item[(ix)] Point symmetry $\kappa(x,y) = \kappa(-x,-y)$ where 
the surface mass profile can be assumed to be symmetric (3+1 quad
configurations and the double \textit{Q0957}).

\end{enumerate}

@@

\begin{table*}
  \begin{center}
    \begin{tabular}{l r r r r l}
      Lens    & \multicolumn{1}{c}{$z_{L}$} & \multicolumn{1}{c}{$z_{S}$} & \multicolumn{1}{c}{$\Delta\theta$ [kpc]} & \multicolumn{1}{c}{$R_{L}$/$R_{e}$} & Environment \\ \hline
      0047 & 0.485 & 3.60 & 12.82 & $1.45\pm0.04$ & G(9) \\
      0414 & 0.960 & 2.64 & 16.01 & $1.85\pm0.05$ & ... \\
      0712 & 0.410 & 1.34 & 6.82  & $1.15\pm0.03$ & ... \\
      0911 & 0.769 & 2.8  & 23.16 & $3.09\pm0.05$ & C \\
      0957 & 0.356 & 1.41 & 29.98 & $3.51\pm0.04$ & C \\
      1115 & 0.310 & 1.72 & 10.76 & $2.94\pm0.06$ & G(13) \\
      1422 & 0.337 & 3.62 & 6.02  & $4.49\pm0.06$ & G(17) \\
      1608 & 0.630 & 1.39 & 13.92 & $1.82\pm0.01$ & !CHECK! \\
      2016 & 1.010 & 3.3  & 26.22 & $6.12\pm0.14$ & C(69) \\
      2045 & 0.870 & 1.28 & 14.46 & $1.48\pm0.03$ & !SEE TEXT! \\
      2237 & 0.039 & 1.7  & 1.40  & $0.89\pm0.01$ & ... \\
    \end{tabular}
    \caption[width=\linewidth]{Table with lens properties \citep[see][]{2011ApJ...740...97L}. References in Section~\ref{sec:data} \textbf{expand}}
    \label{tab:lensproperties}
  \end{center}
\end{table*}

\begin{table*}
  \begin{center}
    \begin{tabular}{l r r r r l l}
      Lens    & \multicolumn{1}{c}{A ['']} & \multicolumn{1}{c}{B ['']} & \multicolumn{1}{c}{C ['']} & \multicolumn{1}{c}{D ['']} & Time delays [days] & Additional priors \\ \hline
      0047 & 1.270 & -0.630 & 0.520 & -0.730 & ... & \\
           & 0.105 & -0.995 & -1.045 & 0.705 & &  \\
      0414 & -0.472 & -1.061 & -1.1947 & 0.885 & ... & \\
           & 1.277 & -0.661 & -0.255 & -0.361 & &  \\
      0712 & -0.013 & 0.795 & 0.747 & -0.391 & ... & \\
           & -0.804 & -0.156 & -0.292 & 0.307 & &  \\
      0712 & -0.013 & 0.795 & 0.747 & -0.391 & ... & \\
           & -0.804 & -0.156 & -0.292 & 0.307 &  &  \\
      0957$^{a}$ & 1.408 & 0.182 & 2.860 & -1.540 & $\Delta t_{BA}=416.5^{+1}_{-1}$ & Symm \\
           & 5.034 & -1.018 & 3.470 & -0.050 &  &  \\
      1115 & 0.355 & -0.909 & -1.093 & 0.717 & $\Delta t_{BA}=12.0^{+2}_{-2} \ \Delta t_{DC}=4.4^{+3.2}_{-2.4}$ & \\
           & 1.322 & -0.714 & -0.260 & -0.627 & &  \\
      1422 & 1.079 & 0.357 & 0.742 & -0.205 & ... & Symm \\
           & -0.095 & 0.973 & 0.656 & -0.147 &  &  \\
      1608 & -1.300 & -0.560 & -1.310 & 0.570 & $\Delta t_{BA}=31.5^{+2}_{-1} \ \Delta t_{CA}=36.0^{+1.5}_{-1.5} \ \Delta t_{DA}=77.0^{+2}_{-1}$ & \\
           & -0.800 & 1.160 & 0.700 & -0.080 &  &  \\
      2016 & -1.735 & 0.335 & 0.437 & 1.268 & ... & \\
           & 1.778 & -1.450 & -1.435 & 0.276 &  &  \\
      2045 & 1.121 & 1.409 & 1.255 & -0.507 & ... & $\gamma(0.1)$, Symm \\
           & 0.824 & 0.035 & 0.576 & -0.183 &  &  \\
      2237 & 0.598 & -0.075 & 0.791 & -0.710 & ... & \\
              & 0.758 & -0.939 & -0.411 & 0.271 &  &  \\
    \end{tabular}
    \caption[width=\linewidth]{Table with lens properties relevant for modelling (positions, time delays) \newline $^{a}$ Images A and B are a pair, C and D the other \textbf{expand}}
    \label{tab:lensmodelling}
  \end{center}
\end{table*}
